SDR enhancement by clouds

### 2024-03-26 article/article.tex
Commit 2024-03-24 21:12
Commit 2024-03-26 19:09

\title{SDR enhancement by clouds}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A Short Subtitle}
\author{Athanasios N. Natsis\footnote{Laboratory of Atmospheric Physics \emph{\href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics \emph{\href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}}}} \and Charikleia Meleti\footnote{Laboratory of Atmospheric Physics \emph{\href{mailto:meleti@auth.gr}{\nolinkurl{meleti@auth.gr}}}}}
\date{2024-03-24}

@@

CE and ECE usefulness
importance
problems

reported trends

CE mechanism

In this study, we define as cloud enhancement events (CE) the cases when the measured
global horizontal irradiance (GHI) at ground level, exceeds the expected value under
clear-sky conditions.
Similar, we define the extreme cloud enhancement events (ECE), the cases when GHI,
exceeds the Total Solar Irradiance on horizontal plane at ground level.
Although the duration of these bursts varies from instantaneous to several minutes,
here we are constrained by the recorded data, to one-minute steps.

Identify CE cases

@@

To be able to detect the CE cases, we had to establish a baseline, above which we can
characterize each data point as an enhancement event, by estimating the occurring
over irradiance (OI). To have an estimation and a first insight on the phenomenon,
we experimented with two simple approaches for the reference. The Haurwitz's model
(Haurwitz 1945), which is as simple clear sky model, and we had already adjusted and
had good fit with our data (Natsis, Bais, and Meleti 2023), and the TSI at the top of the atmosphere.
We have tested both cases by using an appropriate relative threshold and/or an
additional constant offset. The initial results, showed that we can detect a big
portion of the CE events. These results were helpful, and helped to establish some
criteria to further improve the CE identification. It was evident, by inspecting the
daily plot of irradiance, that changes on the atmospheric conditions introduced
numerous false positive and false negative results. To produce a more accurate
reference, we had to take into account more factors that effect the clear sky
radiation. So we used a radiation transfer model.

@@

After evaluating the modeled clear radiation for the different atmospheric
conditions, in relation to the measured GHI data, we choose as a representative of
the clear sky radiation, the case where
\(t_{\text{cs}} = t_{500\text{nm}} - 1\sigma\) and \(w_{h\text{cs}} = w_h - 1\sigma\)
(\(\text{GHI}_\text{mCSi}\)).
These values represent a typical atmosphere in Thessaloniki with low load of aerosols
and humidity, which are the main factor that attenuate the GHI, excluding clouds. To
create the limit of CE identification, we created a two branched threshold, as a
function of SZA. A constant factor for low SZAs, and a higher ramped factor for
higher SZAs (Equation\nobreakspace\ref{eq:CE4}).
This is the criterion of our CE identification.
\begin{equation}
\text{CE} : \begin{cases}
 \text{GHI}_\text{i} > 1.05 \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq 60^\circ$}\\
\text{GHI}_\text{i} > \left ({ 1.18 + \frac{1.05 - 1.18}{60 - 78} \cdot (\theta- 78) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ 78^\circ > \theta > 60^\circ$}\\
\text{Excluded measurements}, & \theta > 78^\circ
\end{cases}\label{eq:CE4}
\end{equation}
where: \(\theta\) is the solar zenith angle, \(\text{GHI}_\text{i}\) the measured
irradiance, and \(\text{GHI}_\text{mCSi}\) the selected modeled clear sky irradiance.

We have to note, that the differentiation of the threshold factor was needed, because
of the high irradiance values we observed, early in the morning and late on the
afternoon. We have confirmed, by inspecting images from the sky cam, that we have
duration of elevated irradiance, either due the clearness of the atmosphere, or some
interferences by reflections on nearby bright surfaces. Although, these cases may
mask some of the CE events. The actual CE events, produce higher irradiance and thus
are identified as such. As a side effect, the reported OI will be slight
underestimated for those SZAs, but the overall contribution of those cases to the
total daily energy, is minimal due the low occurrences and the lower irradiances.

@@

Our dataset, after the data selection processing, consists of
6144534 records of GHI, of which
\(1.799\,\%\) are CE and
\(0.036\,\%\) are ECE events.
The highest GHI recorder was
\(1416.6\,W/m^2\)
on 24~May 2007.
The absolute stronger CE event had an over irradiance of
\(341.79\,W/m^2\) on
15~May 2014.
The relative stronger CE event was
\(54\,\%\) above the
clear sky threshold, on
28~October 2016.

@@

We computed the daily trend of the mean over irradiance of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order autoregressive
model with lag of 1 day, using the `maximum likelihood' fitting method (Gardner, Harvey, and Phillips 1980; Jones 1980) by implementing the function `arima' from the library `stats' of the R
programming language (R Core Team 2023). The trends are reported together with the \(2\sigma\)
error. We observe an increase of
\(0.216\pm 0.083\,W/m^2\)
on the mean over irradiance.

@@

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements. The annual
mean over irradiance is
\(0.23\pm 0.11\,W/m^2\)
(Figure\nobreakspace{}\ref{fig:P-energy-mean}),
but this value alone is not very useful due the intrinsic high variability of this
metric.

@@

\caption{Trends of the mean over irradiance per CE.}\label{fig:P-energy-mean}
\end{figure}

@@

In order to further study the characteristics of the CE events, we grouped the singe
minute CE events, to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can have a duration of multiple of minutes. We have
identified 29009 groups, where the group with the longest duration lasted
140 minutes on
07~July 2013.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude that longer durations are getting increasingly rare, with durations above 10 minutes very rare.

@@

>>>>

\title{SDR enhancement by clouds}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A Short Subtitle}
\author{Athanasios N. Natsis\footnote{Laboratory of Atmospheric Physics \emph{\href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics \emph{\href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}}}} \and Charikleia Meleti\footnote{Laboratory of Atmospheric Physics \emph{\href{mailto:meleti@auth.gr}{\nolinkurl{meleti@auth.gr}}}}}
\date{2024-03-26}

@@

The shortwave solar irradiance, reaches the Earth's surface, is the main energy
source of the planet. It drives and governs the climate.
It has direct practical application, in different energy production method.
\ldots. cite \ldots.
The variability of its intensity, can cause difficulties in predicting the yield and
designing the appropriate equipment.
\ldots. cite \ldots.
An aspect of this variability are the interaction with the clouds, where locally, under some
conditions the total shortwave irradiance reaching the ground can be higher than the
expected clear sky irradiance.
\ldots. cite \ldots.
Those enhancements have been reported to be able to exceed \ldots. and duration.
\ldots. cite \ldots.

The underling mechanism of those events have been studied, with the main factor to be
\ldots. cite \ldots{}
The forward scattering of clouds near the sun has been observed to produce cases of
extreme enhancement irradiances (Yordanov, Saetre, and Midtgård 2015).

Gueymard, 2017; Veerman et al., 2022)
Methods of identification in the literature include the use of a clear sky radiation
modelel

\ldots. cite \ldots.

\ldots{} reported trends ?\ldots{}
\ldots. cite \ldots.

In this study, we used modeled clear sky irradiance, in order to identify cloud
enhancement. We studied their occurrences, intensity and duration, in a thirty year
period at the city of Thessaloniki. We were able to determine some trends of the
phenomenon, it's Climatology, and some other statistical characteristics.

In the relative bibliography different definition have been used for these events, some of
the are summarized by (\textbf{Gueymard2017?}).
Here, we defined as cloud enhancement events (CE) the cases when the measured
global horizontal irradiance (GHI) at ground level, exceeds the expected value under
clear-sky conditions.
Similar, we defined as extreme cloud enhancement events (ECE), the cases when GHI,
exceeds the Total Solar Irradiance on horizontal plane at ground level.
Although the duration of these bursts varies, from instantaneous to several minutes,
here we are constrained by the recorded data, to one-minute steps.

@@

To be able to detect the CE cases, we had to establish a baseline, above which we can
characterize each data point as an enhancement event, by estimating the occurring
over irradiance (OIR). The OIR, here is defined as the irradiance difference of the
measured one-minute \(\text{GHI}_i\) from the CE identification criterion in
Equation\nobreakspace\ref{eq:CE4}
(\(\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}\)).
To have an estimation and a first insight on the phenomenon, we experimented with two
simple approaches for the reference. The Haurwitz's model (Haurwitz 1945), which is
as simple clear sky model, and we had already adjusted and had good fit with our data
(Natsis, Bais, and Meleti 2023), and the TSI at the top of the atmosphere. We have tested both cases
by using an appropriate relative threshold and/or an additional constant offset. The
initial results, showed that we can detect a big portion of the CE events. These
results were helpful, and helped to establish some criteria to further improve the CE
identification. It was evident, by inspecting the daily plot of irradiance, that
changes on the atmospheric conditions introduced numerous false positive and false
negative results. To produce a more accurate reference, we had to take into account
more factors that effect the clear sky radiation. So we used a radiation transfer
model.

@@

After evaluating the modeled clear radiation for the different atmospheric
conditions, in relation to the measured GHI data, we choose as a representative of
the clear sky radiation, the case where
\(t_{\text{cs}} = t_{500\text{nm}} - 1\sigma\) and \(w_{h\text{cs}} = w_h - 1\sigma\)
(\(\text{GHI}_\text{mCSi}\)).
These values represent a typical atmosphere in Thessaloniki with low load of aerosols
and humidity, which are the main factor that attenuate the GHI, excluding clouds. To
create the limit of CE identification, we created a two branched threshold, as a
function of SZA. A constant factor for low SZAs, and a higher ramped factor for
higher SZAs (Equation\nobreakspace\ref{eq:CE4}).
This is the criterion of our CE identification.
\begin{equation}
\text{CE} : \text{GHI}_\text{i} > \text{GHI}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{GHI}_\text{CSlim,i} = 1.05 \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq 60^\circ$}\\
\text{GHI}_\text{CSlim,i} = \left ({ 1.18 + \frac{1.05 - 1.18}{60 - 78} \cdot (\theta- 78) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ 78^\circ > \theta > 60^\circ$}\\
\text{Excluded measurements}, & \theta > 78^\circ
\end{cases}\label{eq:CE4}
\end{equation}
where: \(\theta\) is the solar zenith angle, \(\text{GHI}_\text{i}\) the measured
irradiance, and \(\text{GHI}_\text{mCSi}\) the selected modeled clear sky irradiance.

We have to note, that the differentiation of the threshold factor was needed, because
of the high irradiance values we observed, early in the morning and late on the
afternoon. We have confirmed, by inspecting images from the sky cam, that we have
duration of elevated irradiance, either due the clearness of the atmosphere, or some
interferences by reflections on nearby bright surfaces. Although, these cases may
mask some of the CE events. The actual CE events, produce higher irradiance and thus
are identified as such. As a side effect, the reported OIR will be slight
underestimated for those SZAs, but the overall contribution of those cases to the
total daily energy, is minimal due the low occurrences and the lower irradiances.

@@

Our dataset, after the data selection processing, consists of
6144534 records of GHI, of which
\(1.799\,\%\) are CE and
\(0.036\,\%\) are ECE events.
The highest GHI recorder was
\(1416.6\,W/m^2\)
on 24~May 2007.
The absolute stronger CE event had an OIR of
\(341.79\,W/m^2\) on
15~May 2014.
The relative stronger CE event was
\(54\,\%\) above the
clear sky threshold, on
28~October 2016.

@@

We computed the daily trend of the mean OIR of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order autoregressive
model with lag of 1 day, using the `maximum likelihood' fitting method (Gardner, Harvey, and Phillips 1980; Jones 1980) by implementing the function `arima' from the library `stats' of the R
programming language (R Core Team 2023). The trends are reported together with the \(2\sigma\)
error. We observe an increase of
\(0.216\pm 0.083\,W/m^2\)
on the mean OIR.

@@

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements. The annual
mean OIR is
\(0.23\pm 0.11\,W/m^2\)
(Figure\nobreakspace{}\ref{fig:P-energy-mean}),
but this value alone is not very useful due the intrinsic high variability of this
metric.

@@

\caption{Trends of the mean OIR per CE.}\label{fig:P-energy-mean}
\end{figure}

@@

In order to further study the characteristics of the CE events, we grouped the single
minute CE events, to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can have a duration of multiple of minutes. We have
identified 29009 groups, where the group with the longest duration lasted
140 minutes on
07~July 2013.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude that
longer durations are getting increasingly rare, with durations above 10 minutes very
rare.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Yordanov2015}{}}%
Yordanov, Georgi Hristov, Tor Oskar Saetre, and Ole-Morten Midtgård. 2015. {``Extreme Overirradiance Events in Norway: 1.6 Suns Measured Close to 60°n.''} \emph{Solar Energy} 115 (May): 68--73. \url{https://doi.org/10.1016/j.solener.2015.02.020}.
